### 2024-04-12 paper/paper.qmd
Add files via upload

---
title: "Beyond the Arc: How 3 point shooting has impacted winning in the modern NBA"
subtitle: "Analysing the relationship between a team's 3 point shooting and sucess in the regular season"
author: 
  - Aryaman Suri
thanks: "Code and data are available at: https://github.com/aryamansuri/NBA-3PointAnalysis"
date: today
date-format: long
abstract: "In recent years, the three-point shot has emerged as a pivotal aspect of basketball strategy in the NBA. This study uses statistical modelling tools to analyze the association between three-point shooting and winning in the current NBA. Using season-wise NBA statistics, I examine the cumulative effect of teams' three-point shooting on their overall victory total. This study quantifies this link using Generalized Linear Models, which take into account parameters such as the amount of three-point shots made and other pertinent variables. The findings shed light on the importance of three-point shooting as an NBA success factor. I examine the ramifications of the findings for team strategy, player development, and basketball analytics. Additionally, ethical concerns and study limitations are addressed."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| 
#### Workspace set-up #####

#Load packages
library(tidyverse)
library(dplyr)
library(openxlsx)
library(palmerpenguins)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(ggbeeswarm)
library(ggrepel)

#read in data #
cleaned_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
show_col_types = FALSE
grouped_data <- read.csv(here::here("data/analysis_data/grouped_data.csv"))
show_col_types = FALSE

#read in model #
nba_model <- readRDS(file = here::here("model/nba_model.rds"))
nba_model2 <- readRDS(file = here::here("model/nba_model_team.rds"))
```

# Introduction

The three point line was introduced in the NBA in the 1979-80 season and has brought unprecedented balance to the game. Instead of the game being centered around the big man or athletic wings, the 3 point shot gave everybody on the team a new powerful weapon on offenses. Today, there are 230 players who averaged at least one made three pointer in the 2022-23 regular season. The estimand of this paper is to see if these 3 point shots help the team win games in the regular season.

This paper aims to estimate the association between three-point shooting measures and win probability using advanced statistical techniques after meticulously analyzing season-wise data. The findings highlighted the importance of three-point shooting efficiency as a predictor of team performance in modern basketball and how important it is to be a good three point shooting team for success in today's game.

Understanding this link has far-reaching ramifications for numerous parties, including coaches, players, analysts, and team administration. This study adds to the ongoing debate in the NBA about strategic decision-making and player development by explaining the significance of three-point shooting in deciding game outcomes. Essentially, the question "Does three point shooting actually lead to team success?", is being asked.

I begin by exploring the underlying data, which will provide insights into the variables and metrics used in our analyses. Following that, I explain the complexities of the modelling approach, describing the reasoning behind the statistical techniques and justifying the methodology. Later, the results of the analysis are discussed, diving into the empirical findings and their implications for understanding the relationship between three-point shooting efficiency and NBA team success. Then  I have included a thorough conversation, contextualizing the findings within the larger landscape of basketball analytics and emphasizing significant takeaways from this research. Finally, I summarize the findings and propose paths for future research to further the understanding of this dynamic interaction.

# Data

In this study, I utilized data sourced from the NBA's official database (@cite-NBA) to examine the relationship between various team performance metrics and their corresponding winning percentages throughout the regular season. My analysis aims to uncover insights into the factors contributing to team success in the NBA, specifically 3 point shooting. The NBA's database serves as a comprehensive repository of statistical information on teams, players, and games, providing valuable insights into the dynamics of professional basketball. Below, I provide a detailed overview of the data collection process, cleaning procedures, and subsequent analysis.

## Data Collection

All raw data utilized in this study was collected from the official NBA database (@cite-NBA), which aggregates comprehensive statistics on team performance, player performances, and game outcomes throughout the league's history. Since this study only focuses on the modern NBA, which is defined as the era from the 2014-15 season onward, the season wise data had to be individually downloaded and combined. Additionally, since the 2019-20 season was shortened due to the COVID-19 outbreak and an equal number of games was not played by all teams, we will not be considering that season for the sake of this study. This data is then stored in our raw_data2.xlsx file using @cite-writexl. This dataset comprises a wide array of variables, including team statistics such as points scored, rebounds, assists, turnovers, and various other performance metrics recorded over multiple seasons.

## Data Cleaning

Upon acquiring the raw data from nba.com (@cite-NBA), extensive cleaning procedures were implemented to ensure its suitability for analysis. This involved addressing missing or erroneous values, standardizing variable names, and aggregating data at the team level to facilitate meaningful comparisons across different seasons and teams. Additionally, I computed additional variables such as win percentages, and filtered the data to only include winning stats and 3 point stats using the *group_by* (@cite-groupby) package. This dataset is then stored in the analysis data as a csv file (@cite-writecsv).

## Data Analysis

With the cleaned dataset in hand as seen in @tbl-1, I conducted a rigorous analysis to explore the relationships between 3 point shot performance metrics and team success in the NBA. Various statistical techniques were employed, including descriptive analysis, correlation analysis, and regression modeling, to identify significant predictors of winning percentages. Furthermore, I visualized the data through graphs and tables to elucidate trends and patterns in team performance across different seasons.

```{r}
#| echo: false
#| eval: true
#| label: tbl-1
#| tbl-cap: "Overview of the cleaned dataset"
#| warning: false

a <- head(cleaned_data)
colnames(a) <- c('Team','Games', 'Wins', 'Loss', 'Win%', '3PM', '3PA', '3P%', 'Box +/-')
kable(a)
```

By segmenting the data by season and team, I was able to conduct a detailed examination of performance trends over time and assess the impact of various factors on team success. This analysis provides valuable insights for coaches, analysts, and basketball enthusiasts seeking to understand the drivers of success in the NBA.

## Variables of interest

In this study, our focus is solely on win counts and three-point statistics. With this objective in mind, our key variables of interest include Games Played (**GP**), Wins (**W**), Losses (**L**), Three-Pointers Made per game (**3PM**), Three-Pointers Attempted per game (3PA), and Three-Point Shooting Percentage (**3P%**). Our modeling approach will specifically incorporate Wins (**W**) and **3PM** to explore their relationship.

While the measurement of this model's data can be criticized by NBA fans , we've deliberately opted for **3PM** over **3PA** and **3P%** for several reasons. **3PM** serves as a robust metric, capturing a team's volume of successful three-point shots with reasonable efficiency. Unlike **3PA**, which might bias the model towards teams overly reliant on three-point attempts irrespective of their success rate, **3PM** provides a balanced representation of both volume and efficiency in three-point shooting. Similarly, **3P%** could skew the analysis towards teams with lower volumes of three-point attempts, overlooking the broader spectrum of team performance.

By prioritizing **3PM** alongside Wins, our model aims to uncover nuanced insights into the impact of three-point scoring on overall team success while mitigating potential biases associated with alternative metrics. This strategic selection ensures a comprehensive exploration of the relationship between three-point performance and winning outcomes in NBA games.

# Model

The goal of my modeling strategy is two fold. First, I want to analyze the effect of 3 point shooting on team success defined by wins. Secondly, I want to predict NBA team wins based on the number of 3-point field goals made by each team throughout the regular season.

Here I briefly describe the Bayesian analysis model used to investigate the relationship between 3-point field goals made and team wins.

## Model set-up

Let $y_i$ represent the number of wins for team $i$ throughout a given season, and let $\beta_{\text{3pts}}$ denote the effect of 3-point field goals made on team wins. The number of 3-point field goals made per game ($\text{3pts}_{i}$) by team $i$ is the predictor variable considered in this model.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_{\text{3pts}} \times \text{3pts}_{i} \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_{\text{3pts}} &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@cite-R] using the `rstanarm` (@cite-rstanarm) package. The default priors from `rstanarm` are utilized.

## Model justification

I hypothesize a positive relationship between the number of 3-point field goals made by a team and their number of wins throughout the regular NBA season. This hypothesis is rooted in the understanding that teams with higher 3-point shooting efficiency typically demonstrate stronger offensive capabilities, which often translate into higher-scoring games and, consequently, more victories. This relationship is particularly pertinent in today's NBA landscape, where teams increasingly prioritize offensive strategies and exploit the scoring potential offered by 3-point shots.

The chosen Bayesian analysis model (@cite-rstanarm) aims to quantitatively assess the impact of 3-point field goals made on team wins and to provide insights into the significance of this performance metric in determining overall team success. By modeling the relationship between 3-point field goals made and wins using a normal distribution for the response variable, we acknowledge the continuous nature of team wins and seek to capture the variability in win counts across different levels of 3-point shooting proficiency. This approach allows for a nuanced understanding of how variations in 3-point performance contribute to variations in team outcomes, thereby elucidating the importance of 3-point shooting in the context of NBA team success.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Team success based on 3 Points"
#| warning: false

library(knitr)

# Model summary statistics
model_summary <- data.frame(
  Parameters = c("(Intercept)", "3PM", "sigma", "mean_PPD"),
  Mean = c(27.8, 1.2, 11.6, 40.3),
  SD = c(3.5, 0.3, 0.5, 1.1),
  P10 = c(23.4, 0.8, 10.9, 39.0),
  P50 = c(27.8, 1.2, 11.5, 40.3),
  P90 = c(32.2, 1.6, 12.3, 41.7),
  mcse = c(0.1, 0.0, 0.0, 0.0),
  Rhat = c(1.0, 1.0, 1.0, 1.0),
  n_eff = c(3811, 3848, 3536, 3764)
)

# Print table
kable(model_summary, 
      format = "markdown")
```

In the context of the NBA, the interpretation of the model is provided below.

The **Intercept (27.8)** represents the estimated mean number of wins when the number of 3-point shots made (3PM) is zero. In other words, when a team makes zero 3-point shots, the model estimates that they would win approximately 27.8 games on average. The **3PM (1.2)** coefficient represents the change in the estimated mean number of wins for each additional 3-point shot made by a team. In this case, for every additional 3-point shot made per game over the season, the model predicts an increase of approximately 1.2 wins on average.
The **Sigma (11.6)** parameter represents the estimated standard deviation of the error term in the model. It indicates the variability in team success (measured by the number of wins) that is not explained by the number of 3-point shots made. The **Mean_PPD (40.3)**represents the average posterior predictive distribution of the outcome variable (number of wins). It provides an estimate of the average number of wins predicted by the model.

Overall, based on these estimates, it appears that there is a positive association between the number of 3-point shots made by a team and their success, as measured by the number of wins. Teams that make more 3-point shots tend to win more games on average, according to the model. However, it's essential to consider other factors not included in the model that may also influence team success.

```{r}
#| echo: false
#| label: fig-2
#| fig-cap: "Graph depicting the **Wins** vs the **3 Points Made** per game"
#| warning: false
ggplot(cleaned_data, aes(x = X3PM, y = W)) +
  geom_point(color = "orange") +
  labs(x = "3PM", y = "Wins", title = "Wins vs 3PM") +
  theme_minimal() +  # Apply a minimal theme
  theme(
    axis.line = element_line(color = "black"),  # Add axis lines
    panel.grid.major = element_blank(),         # Remove major gridlines
    panel.grid.minor = element_blank()          # Remove minor gridlines
  )
```

@fig-2 shows us a visual representation of how 3 point shots made are related to the wins of a team. The graph illustrates a positive correlation between the number of 3-point shots made per game and team wins, indicating the importance of proficient long-range shooting in securing victories. However, outliers exist, suggesting that factors beyond 3-point shooting also influence team success. We discuss these in detail in the discussion portion.

If we group our statistics by team (@cite-groupby), we can gain valuable insights into which teams benefit most from their three-point shooting prowess. @tbl-3 presents a detailed breakdown of win statistics and three-point shooting statistics grouped by team for the seasons of interest. The table is sorted in descending order of wins, allowing us to identify the top-performing teams. By examining this table, we can discern trends, such as the correlation between three-point shooting and team success, and identify standout performers in each category.

```{r}
#| echo: false
#| eval: true
#| label: tbl-3
#| tbl-cap: "Table showing the win stats and 3 point stats grouped by the Team"
#| warning: false

a <- grouped_data
colnames(a) <- c('Team', 'Games', 'Wins','3PM', '3PA', 'Avg Wins', 'Avg 3PM')
sorted_table <- a %>%
  arrange(desc(Wins))

# Display the sorted table
kable(sorted_table)
```

```{r}
#| echo: false
#| label: fig-1
#| fig-cap: "Graph depicting the **Average team wins** and **predicted team wins** vs **Average three point shots made** per game"
#| warning: false
# Calculate predicted wins
predicted_wins <- fitted(nba_model2)

# Create the plot
ggplot(grouped_data, aes(x = Average_3PM, y = Average_Wins)) +
  geom_point(color = "black") +
  geom_line(aes(y = predicted_wins, color = "Predicted Wins")) +
  labs(x = "Average 3PM", y = "Average Wins", title = "Average Wins vs. Average 3PM") +
  theme_minimal() +  # Apply a minimal theme
  theme(
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    axis.line = element_line(colour = "black"),  # Set color of axis lines
    text = element_text(size = 12),  # Adjust text size
    plot.title = element_text(hjust = 0.5),  # Center the plot title
    legend.position = "top"  # Set legend position
  ) +
  scale_color_manual(values = "orange", name = "")  # Set legend title and color
```

@fig-1, plotted using ggplot (@cite-ggplot) illustrates the association between the average number of 3-point shots made per game and the average number of wins per season for NBA teams. The black points represent observed average wins, while the orange line depicts predicted wins based on a statistical model. This model suggests a relationship between 3-point shooting proficiency and team success. Comparing the orange line with the black points allows for an evaluation of the model's predictive accuracy.

# Discussion

In this study, we aimed to investigate the relationship between three-point shooting efficiency and team success in the NBA. Our analysis involved a comprehensive examination of season-wise NBA statistics, focusing on variables such as three-pointers made (3PM) and team wins (W). We employed advanced statistical techniques, including data cleaning, correlation analysis, and regression modeling, to uncover insights into the impact of three-point shooting on team performance. By analyzing the data and modeling the relationship between 3PM and W, we sought to shed light on the significance of three-point shooting as a predictor of NBA team success.

Our analysis revealed a compelling association between three-point shooting proficiency and team success in the NBA. We found that teams with higher rates of three-pointers made per game tended to achieve more wins over the course of the regular season. This suggests that three-point shooting plays a crucial role in determining the outcome of NBA games, with proficient long-range shooters contributing significantly to their team's overall success. These findings underscore the strategic importance of three-point shooting in modern basketball and highlight its impact on team performance.
Additionally, our analysis uncovered nuanced insights into the dynamics of three-point shooting and team success. We observed variations in the relationship between 3PM and W across different teams and seasons, indicating that the impact of three-point shooting may vary based on contextual factors such as team composition, playing style, and opponent strategies. Furthermore, our findings suggest that while three-point shooting is a significant predictor of team success, it is not the sole determinant, with other factors such as defense, rebounding, and overall offensive efficiency also playing crucial roles in determining game outcomes.

## Weaknesses

Despite the rigor of our analysis, several limitations warrant consideration. One potential limitation is the reliance on aggregate team statistics, which may overlook individual player contributions and team dynamics that could influence the relationship between 3PM and W. Additionally, our analysis focused solely on regular-season data, excluding playoff games, which may have different dynamics and performance factors since the game is known to "slow down" in the playoffs. This can be caused by teams specifically executing strategies to identify and take advantage of the opponents weaknesses over a 7 game series.

Furthermore, while our modeling approach accounted for various factors, including team performance metrics and game outcomes, it may not capture all relevant variables that could impact the relationship between three-point shooting and team success. Success in the NBA cannot simply be boiled down to efficient volume three point shooting since it is a nuanced game with many other factors contributing to a team's overall success.

## Drawbacks of the 3 point revolution

Although, this paper discusses the upside of the three point line, I feel it is important to also include it's drawbacks. An interesting example of this in the modern NBA is the 2017-18 Houston Rockets. The Houston Rockets' general manager was a huge fan of the three point shot and in the 2017-18 season built an entire roster around 3 point shooting (@cite-Rockets1). This team was a huge success in the regular season and they shot more 3 pointers than any other team in league history. What he failed to take into account was, that basketball isn't math. While the Rockets were the most successful regular season team with 65 wins (@cite-Rockets2), they matched up against the Golden State Warriors in the western conference finals of the playoffs. The series went to seven games, and in game 7, the Rockets essentially shot themselves out of the game. They went 7 for 49 from the arc as a team, which was a record for the worst 3 point shooting in NBA history. This team lived by the three and eventually would die by the three.

In essence, while the three-point shot remains a potent weapon in the modern NBA, its effectiveness is contingent upon a nuanced understanding of its role within a team's overall offensive strategy. As the Houston Rockets learned, basketball is not solely a game of statistics—factors such as matchup dynamics, defensive schemes, and clutch performances can influence outcomes in ways that transcend numerical analysis.

## Future Research

Looking ahead, there are plenty of opportunities for future research to build upon our discoveries and tackle the limitations of our study. One promising avenue is to delve deeper into the data by analyzing individual player-level statistics. This approach could help us better grasp how specific players contribute to their team's success, providing a more nuanced understanding of the role of each player in three-point shooting dynamics.

Moreover, conducting longitudinal studies that span multiple seasons could offer valuable insights into the evolving trends and patterns of three-point shooting in the NBA. By tracking team performance over time, researchers could uncover long-term dynamics and factors that influence the effectiveness of three-point strategies.

Additionally, qualitative research methods, such as interviews with players and coaches, present an opportunity to gain unique perspectives on the strategic significance of three-point shooting. Engaging directly with those involved in the game could offer valuable insights into the decision-making processes and team dynamics surrounding three-point strategies.

By exploring these avenues of inquiry, future research has the potential to deepen our understanding of the intricate relationship between three-point shooting and team success in the NBA, ultimately contributing to advancements in basketball analytics and strategy.

# References